Systemix MAG-SMAG Profiling

### 2024-07-04 R/paper.Rmd
added latest changes

---
title: "Systemix Article: Combined MAG-SMAG Profiling"
author: "Moises A. Rojas"
date: "2023-12-01 // mod. Mar 2024"
output: github_document
---

@@

```{r magdata}
# Read Systemix MAGs taxonomy and QC files
gtdb_taxonomy <- read.csv("1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)
mags_qc <- read.csv("2-statistics/checkm_parsed_bins.tsv", sep = "\t")

@@

# Define final attributes
mag_samples['mag_num'] <- nrow(mag_samples)
mag_samples['ecosystem'] <- "Biomining"
mag_samples['continent'] <- "South America"

@@

# Rename columns
colnames(smag_data)[6] <- "genome_size"
colnames(smag_data)[8] <- "kingdom"
colnames(smag_data)[16] <- "ecosystem"
colnames(smag_data)[17] <- "continent"

colnames(smag_data)[7] <- "kingdom"
colnames(smag_data)[14] <- "ecosystem"
colnames(smag_data)[15] <- "continent"

@@

## Section 1: In-depth taxonomy analysis

### Biomining MAGs vs SMAG taxa comparison

@@

# Keep only bacteria and defined ecosystems (discard archaea kingdom and non-NA eco.)
mags_metadata_bac <- mags_metadata %>%
  dplyr::filter(kingdom == "Bacteria" & !is.na(ecosystem)) #n=32975

@@

plot_depth_top_phyla <- ggplot(mags_md_top_phy, aes(x = freq, y = reorder(phy_top, freq))) +
  geom_bar(stat = "identity", fill = "#459682") +
  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.5, size = 2.5) +
  labs(x = "Number", y = "Phylum") +
  scale_x_continuous(limits = c(0, 10000)) +
  theme_classic()
ggsave("plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 7, width = 6)

@@

plot_depth_spec <- ggplot(mags_depth_spec) +
  geom_bar(aes(x = species, fill = phylum)) +
  labs(x = "", y = "Number of kSGBs from biomining") +
  scale_fill_manual(values = "#78Acc2") +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = c(0.88, 0.12),
        strip.background = element_blank()) +#axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1)
  facet_grid(.~ecosystem) +
  coord_flip()
ggsave("plot_md_biom_species.pdf", plot_depth_spec, height = 2.5, width = 8)

@@

plot_depth_spec_eco <- ggplot(mags_depth_spec_eco, aes(x = reorder(ecosystem, ratio),
                                                       y = ratio, fill = ecosystem)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.5, size = 3) +
  labs(x = "Ecosystem", y = "Number of kSGBs / MAGs") +
  scale_y_continuous(limits = c(0, 10)) +
  scale_fill_manual(values = mags_colors_eco) +
  theme_classic() +
  theme(legend.position = "none") +
  coord_flip()
ggsave("plot_md_species_eco.pdf", plot_depth_spec_eco, height = 7, width = 6)

@@

plot_depth_sgbs <- ggplot(mags_depth_sgbs[1:60, ],
                          aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
  scale_y_continuous(limits = c(0, 8000)) +
  labs(x = "Phylum", y = "Number of species") +
  theme_classic() +
  theme(legend.position = c(0.85, 0.15), strip.placement = "outside") +
  coord_flip()
#ggsave("plot_md_SGBs_phylum.pdf", plot_depth_sgbs, height = 7, width = 6)

@@

plot_depth_sgbs_p <- ggplot(mags_depth_sgbs_p[1:60, ],
                            aes(x = reorder(phylum, t_phy), y = tot_p, fill = cond_p)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "", y = "% of species") +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank()) +
  coord_flip()
#ggsave("plot_md_SGBs_phylum_p.pdf", plot_depth_sgbs_p, height = 7, width = 5)

plot_depth_sgbs_both <- (plot_depth_sgbs | plot_depth_sgbs_p) +
  plot_layout(nrow = 1, ncol = 2, widths = c(0.55, 0.5))
ggsave("plot_md_SGBs_phylum.pdf", plot_depth_sgbs_both, height = 7, width = 7)

@@

plot_depth_sgbs_biom <- ggplot(mags_depth_sgbs_biom,
                               aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#DD7E8B", "#459682")) +
  scale_y_continuous(limits = c(0, 25)) +
  labs(x = "Phylum", y = "Biomining species") +
  theme_classic() +
  theme(legend.position = c(0.75, 0.2), strip.placement = "outside") +
  coord_flip()
ggsave("plot_md_SGBs_phylum_biom.pdf", plot_depth_sgbs_biom, height = 5, width = 4)

@@

plot_depth_usgbs <- ggplot(mags_depth_usgbs_eco,
                           aes(x = num, y = reorder(phylum, num), fill = ecosystem)) +
  geom_col() +
  #geom_text(aes(label = num), vjust = 0.5, hjust = -0.5, size = 2.5) +
  labs(x = "Number of uSGBs", y = "Phylum") +
  scale_x_continuous(n.breaks = 3) +
  scale_fill_manual(values = mags_colors_eco) +
  theme_classic() +
  theme(legend.position = "none",
        strip.placement = "outside",
        strip.background = element_blank(),
        panel.border = element_blank()) +
  facet_grid(.~ecosystem, scales = "free")
ggsave("plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 8, width = 12)

@@

plot_depth_kggbs_biom <- ggplot(mags_depth_kggbs_biom,
                                aes(y = reorder(genus, norm), x = norm, fill = ecosystem)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_x_continuous(limits = c(0, 101), n.breaks = 5) + #check!
  scale_fill_manual(values = mags_colors_eco) +
  labs(x = "% kGGBs", y = "Known genus from biomining") +
  theme_classic() +
  theme(legend.title = element_blank())
ggsave("plot_md_kGGBs_biom.pdf", plot_depth_kggbs_biom, height = 7, width = 6)

@@

```{r magbiomqc}
# Read QC file from CheckM
mags_qc <- read.csv("mags-all/2-statistics/checkm_parsed_bins.tsv", sep = "\t")

@@

# Plot
plot_mags_qc <- ggplot(mags_qcs, aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
  geom_col(position = "dodge2") +
  scale_fill_manual(values = c("#DB7971", "#5CA650", "#859AC7")) +
  labs(x = "Number of MAGs", y = "") +
  theme_classic() +
  theme(legend.position = "none")
ggsave("mags-all/2-statistics/plots/plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 3, width = 5)
```

@@

```{r magbiomtree}
# Load GTDB taxonomy
gtdb_taxonomy <- read.csv("1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)

@@

# Read full tree (S15)
tree <- ape::read.tree("1-taxonomy/trees-gtdbtk/gtdbtk.DASTool-S15.tree")

# Subset to biomining placements
tree_s15 <- keep.tip(tree, as.character(gtdb_taxonomy$genome))
tree_s15 #44 tips

# Export tree to file
if (file.exists("1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree") == FALSE) {
  ape::write.tree(tree_s15, "1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree")
}
rm(tree) #clean memory

# Rename column {genome -> label} (to match with tree) from GTDB data
colnames(gtdb_taxonomy)[1] <- "label"

# Convert subset tree to dataframe and join some taxa
tree_s15_df <- tidytree::as_tibble(tree_s15)
tree_s15_df <- treeio::full_join(tree_s15_df, gtdb_taxonomy[c("label", "phylum", "family", "genus", "species")], by = "label")

@@

# Plot circular tree
plot_tree_s15 <- ggtree(tree_s15_md, layout = "circular", size = 0.2)

# Plot tree adding phylum
plot_tree_s15_phyla <- plot_tree_s15 +
  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1) +
  geom_tiplab(aes(label = phylum, color = phylum), size = 1.75, hjust = -0.1) +
  scale_colour_discrete(breaks = c("c", "d"), na.value = NA) +
  scale_colour_manual(values = tree_colors, na.translate = FALSE, guide = "none") +
  #guides(color = guide_legend(title = "Phyla", ncol = 1, order = 1)) +
  theme()
ggsave("plot_biom_MAGs_tree_phyla.pdf", plot_tree_s15_phyla, height = 5, width = 7)

# Process QC attributes to match tree tips order
tree_tips <- tree_s15$tip.label
tree_com <- vector()
tree_con <- vector()
tree_st  <- vector()
tree_gs  <- vector()
tree_gc  <- vector()
tree_pg  <- vector()

# Match datasets
for (i in 1:nrow(tree_s15_df)) {
  for (j in 1:nrow(mags_qc)) {
    if (tree_s15_df$label[i] == mags_qc$bin[j]) {
      tree_com <- append(tree_com, mags_qc$completeness[j])
      tree_con <- append(tree_con, mags_qc$contamination[j])
      tree_st  <- append(tree_st,  mags_qc$strain_heterogeneity[j])
      tree_gs  <- append(tree_gs,  mags_qc$genome_size[j])
      tree_gc  <- append(tree_gc,  mags_qc$gc[j])
      tree_pg  <- append(tree_pg,  mags_qc$predicted_genes[j])
    }
  }
}

# Save attributes in a dataframe
tree_qc_df <- tibble(Bin = tree_tips, Completeness = tree_com,
                     Contamination = tree_con, Strain_Het = tree_st,
                     Genome_Size = tree_gs, GC = tree_gc, Pred_Genes = tree_pg)
rm(tree_tips, tree_com, tree_con, tree_st, tree_gs, tree_gc, tree_pg) #clean memory

# Plot tree adding QC as multiple layers
plot_tree_s15_md <- plot_tree_s15_phyla +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Completeness, fill = Completeness),
             orientation = "y", stat = "identity", position = "auto", width = 0.85,
             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.6,
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3,
                                title = "Complet.", title.size = 1.6,
                                title.height = 0.035, title.angle = 30)) +
  scale_fill_gradient(low = "#1F4060", high = "#4CA8F6", guide = "none") +
                      #guide = guide_legend(keyheight = 0.4, keywidth = 0.7,
                      #                     reverse = TRUE, order = 1)) +
  ###Error in scale$guide == "none"
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Contamination, fill = Contamination),
             orientation = "y", stat = "identity", position = "auto", width = 0.85,
             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1,
                                title = "Contam.", title.size = 1.6,
                                title.height = 0.035, title.angle = 30)) +
  scale_fill_gradient(low = "#6A60A8", high = "#C4C2DE", guide = "none") +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Strain_Het, fill = Strain_Het),
             orientation = "y", stat = "identity", position = "auto", width = 0.85,
             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3,
                                title = "Strain Het.", title.size = 1.6,
                                title.height = 0.035, title.angle = 30)) +
  scale_fill_gradient(low = "#C9543F", high = "#E9BAB2",
                      name = "Strain Heterogeneity", guide = "none") +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Pred_Genes, fill = Pred_Genes),
             orientation = "y", stat = "identity", position = "auto", width = 0.85,
             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.2, offset = 0.1,
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 1,
                                title = "Pred. genes", title.size = 1.6,
                                title.height = 0.035, title.angle = 30)) +
  scale_fill_gradient(low = "#279854", high = "#A9DEDE",
                      name = "Predicted Genes", guide = "none") +
  theme()
  #theme(legend.title = element_text(size = 8),
  #      legend.text = element_text(size = 7),
  #      legend.key.height = unit(3.5, units = "mm"),
  #      legend.key.width = unit(2, units = "mm"))
ggsave("plot_biom_MAGs_tree_md.pdf", plot_tree_s15_md, height = 7, width = 7) #x9
```

@@

```{r fig1}
# Merge plots using patchwork
fig_1a <- { plot_tree_s15_md | plot_depth_spec_eco +
  plot_layout(ncol = 2, widths = c(1, 0.5))
  } + plot_annotation(tag_levels = list(c("a", "e"))) &
  theme(plot.tag =  element_text(face = "bold"))
ggsave("fig_1a.pdf", fig_1a, height = 7, width = 14)

fig_1b <- { plot_mags_qc / plot_depth_sgbs_biom |
    plot_depth_sgbs_both | plot_depth_kggbs_biom
  } + plot_layout(ncol = 3, widths = c(0.5, 2, 0.75)) +
  plot_annotation(tag_levels = list(c("b", "c", "d", "", "f"))) &
  theme(plot.tag =  element_text(face = "bold"))
ggsave("fig_1b.pdf", fig_1b, height = 7, width = 14)

fig_1 <- (fig_1a / fig_1b) +
  plot_annotation(tag_levels = list(c("a", "e", "b", "c", "d", "", "f"))) &
  theme(plot.tag =  element_text(face = "bold"))
ggsave("fig_1.pdf", fig_1, height = 14, width = 14)
#ggsave("fig_1.png", fig_1, height = 14, width = 14)
rm(fig_1a, fig_1b) #clean memory

#sup 1
fig_s1 <- { plot_depth_top_phyla |
    (plot_depth_usgbs / plot_depth_spec + plot_layout(heights = c(0.85, 0.15))) } +
  plot_layout(ncol = 2, widths = c(0.25, 1)) +
  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
  theme(plot.tag =  element_text(face = "bold"))
fig_s1[[2]][[1]] <- fig_s1[[2]][[1]] + theme(axis.title.y = element_text(vjust = 0))
#fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.text.y = element_text(vjust = 0))
#fig_s1 <- wrap_plots(plot_depth_usgbs, plot_depth_spec, nrow = 2, ncol = 1, heights = c(0.85, 0.15))
ggsave("mags-all/fig_s1.pdf", fig_s1, height = 9, width = 18)
#ggsave("mags-all/fig_s1.png", fig_s1, height = 9, width = 18)
```

@@

# Populate both pathways and metabolism
kegg_ko_paths <- vector()
kegg_ko_metab <- vector()
for (i in 1:nrow(kegg_paths)) {
	kegg_ko_paths <- append(kegg_ko_paths, fx_setKOpathways(kegg_paths$Definition[i]))
	kegg_ko_metab <- append(kegg_ko_metab, fx_setKOmetabtypes(kegg_paths$Definition[i]))
}
kegg_paths['Pathway'] <- kegg_ko_paths
kegg_paths['Metabolism'] <- kegg_ko_metab
rm(kegg_ko_paths, kegg_ko_metab) #clean memory

@@

### Bubble and box plots

>>>>

---
title: "Systemix MAG-SMAG Profiling"
author: "Moises A. Rojas"
date: "2023-12-01 // mod. Mar 2024 // rev. Jun 2024"
output: github_document
---

@@

```{r magdata}
# Read Systemix MAGs taxonomy and QC files
gtdb_taxonomy <- read.csv("article_profiling/1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)
mags_qc <- read.csv("article_profiling/2-statistics/checkm_parsed_bins.tsv", sep = "\t")

@@

# Define final attributes
mag_samples["mag_num"] <- nrow(mag_samples)
mag_samples["ecosystem"] <- "Biomining"
mag_samples["continent"] <- "South America"

@@

# Rename columns
colnames(smag_data)[6] <- "genome_size"
colnames(smag_data)[8] <- "kingdom"
colnames(smag_data)[16] <- "ecosystem"
colnames(smag_data)[17] <- "continent"
colnames(smag_data)[7] <- "kingdom"
colnames(smag_data)[14] <- "ecosystem"
colnames(smag_data)[15] <- "continent"

@@

## Section 1: Taxonomic analysis

### In-depth biomining MAGs vs SMAG comparison

@@

# Keep only bacteria and known ecosystems (n = 32975)
mags_metadata_bac <- mags_metadata %>% dplyr::filter(kingdom == "Bacteria" & !is.na(ecosystem))

@@

plot_depth_top_phyla <- mags_md_top_phy %>%
  ggplot(aes(x = freq, y = reorder(phy_top, freq))) +
  geom_bar(stat = "identity", fill = "#459682") +
  geom_text(aes(label = freq), vjust = 0.5, hjust = -0.25, size = 3) +
  labs(x = "Number", y = "Phylum") +
  scale_x_continuous(limits = c(0, 10000)) +
  theme_classic() +
  theme(axis.text = element_text(size = 14),
        axis.title = element_text(size = 16))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_top_phyla.pdf", plot_depth_top_phyla, height = 7, width = 6)

@@

plot_depth_spec <- ggplot(mags_depth_spec) +
  geom_bar(aes(x = species, fill = phylum)) +
  labs(x = "", y = "Number of kSGBs from biomining") +
  scale_fill_manual(values = "#78acc2") +
  coord_flip() +
  theme_classic() +
  theme(legend.title = element_blank(),
        legend.position = c(0.88, 0.12),
        strip.background = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14),
        strip.text.x = element_text(size = 10)) +
  facet_grid(.~ecosystem)
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_biom_species.pdf", plot_depth_spec, height = 3, width = 9)

@@

plot_depth_spec_eco <- mags_depth_spec_eco %>%
  ggplot(aes(x = reorder(ecosystem, ratio), y = ratio, fill = ecosystem)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(ratio, 2)), vjust = 0.5, hjust = -0.25, size = 3) +
  labs(x = "Ecosystem", y = "Number of kSGBs/MAGs") +
  scale_y_continuous(limits = c(0, 10)) +
  scale_fill_manual(values = mags_colors_eco) +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_species_eco.pdf", plot_depth_spec_eco, height = 7, width = 6)

@@

plot_depth_sgbs <- mags_depth_sgbs[1:60, ] %>%
  ggplot(aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 8000)) +
  labs(x = "Phylum", y = "Number of species") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = c(0.85, 0.15),
        strip.placement = "outside",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

@@

plot_depth_sgbs_p <- mags_depth_sgbs_p[1:60, ] %>%
  ggplot(aes(x = reorder(phylum, t_phy), y = tot_p, fill = cond_p)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 100)) +
  labs(x = "", y = "% of species") +
  coord_flip() +
  theme_classic() +
  theme(legend.position = "none",
        axis.text.y = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))

#merge
plot_depth_sgbs_both <- wrap_plots(plot_depth_sgbs, plot_depth_sgbs_p,
                                   nrow = 1, ncol = 2, widths = c(0.55, 0.5))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_SGBs_phylum.pdf", plot_depth_sgbs_both, height = 7, width = 7)

@@

plot_depth_sgbs_biom <- mags_depth_sgbs_biom %>%
  ggplot(aes(x = reorder(phylum, tot), y = tot, fill = cond)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_fill_manual(name = "", labels = c("kSGB", "uSGB"), values = c("#dd7e8b", "#459682")) +
  scale_y_continuous(limits = c(0, 25)) +
  labs(x = "Phylum", y = "Biomining species") +
  coord_flip() +
  theme_classic() +
  theme(aspect.ratio = 0.875,
        legend.position = c(0.75, 0.3),
        strip.placement = "outside",
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_SGBs_phylum_biom.pdf", plot_depth_sgbs_biom, height = 4, width = 6)

@@

plot_depth_usgbs <- mags_depth_usgbs_eco %>%
  ggplot(aes(x = num, y = reorder(phylum, num), fill = ecosystem)) +
  geom_col() +
  #geom_text(aes(label = num), vjust = 0.5, hjust = -0.5, size = 2.5) +
  labs(x = "Number of uSGBs", y = "Phylum") +
  scale_x_continuous(n.breaks = 3) +
  scale_fill_manual(values = mags_colors_eco) +
  theme_classic() +
  theme(legend.position = "none",
        strip.placement = "outside",
        strip.background = element_blank(),
        panel.border = element_blank(),
        axis.text = element_text(size = 14),
        axis.title = element_text(size = 16),
        strip.text.x = element_text(size = 10)) +
  facet_grid(.~ecosystem, scales = "free")
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_uSGBs_eco.pdf", plot_depth_usgbs, height = 9, width = 14)

@@

plot_depth_kggbs_biom <- mags_depth_kggbs_biom %>%
  ggplot(aes(y = reorder(genus, norm), x = norm, fill = ecosystem)) +
  geom_col(position = position_stack(reverse = TRUE)) +
  scale_x_continuous(limits = c(0, 101), n.breaks = 5) +
  scale_fill_manual(values = mags_colors_eco) +
  labs(x = "% kGGBs", y = "Known genus from biomining") +
  theme_classic() +
  theme(legend.title = element_blank(),
        axis.text = element_text(size = 12),
        axis.title = element_text(size = 14))
ggsave("article_profiling/4-mags-smag/1-plots-taxonomy/plot_md_kGGBs_biom.pdf", plot_depth_kggbs_biom, height = 7, width = 6)

@@

```{r magbiomqc}
# Read QC file from CheckM
mags_qc <- read.csv("article_profiling/2-statistics/checkm_parsed_bins.tsv", sep = "\t")

@@

# Plot
plot_mags_qc <- mags_qcs %>%
  ggplot(aes(x = mags, y = fct_reorder(label, mags), fill = label)) +
  geom_col(position = "dodge2") +
  #geom_text(aes(label = mags), vjust = 0.5, hjust = -0.25, size = 3) +
  scale_fill_manual(values = c("#db7971", "#5ca650", "#859ac7")) +
  labs(x = "Number of MAGs", y = "") +
  theme_classic() +
  theme(aspect.ratio = 0.875,
        axis.text = element_text(size = 12),
        axis.title.x = element_text(size = 14),
        legend.position = "none")
ggsave("article_profiling/2-statistics/plots/plot_biom_MAGs_QC.pdf", plot_mags_qc, height = 4, width = 6)
```

@@

```{r magbiomtree}
# Load GTDB taxonomy
gtdb_taxonomy <- read.csv("article_profiling/1-taxonomy/gtdbtk_classification.tsv", sep = "\t", stringsAsFactors = FALSE)

@@

# Read tree (S15 sample), subset it to biomining placements and save to file
tree <- ape::read.tree("article_profiling/1-taxonomy/raw/gtdbtk-trees/gtdbtk.DASTool-S15.tree")
tree_s15 <- keep.tip(tree, as.character(gtdb_taxonomy$genome))
tree_s15 #44 tips
if (file.exists("article_profiling/1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree") == FALSE) {
  ape::write.tree(tree_s15, "article_profiling/1-taxonomy/trees-subset/gtdbtk.DASTool-S15.tree")
}
rm(tree) #clean memory

# Rename column from GTDB data to match with the tree
colnames(gtdb_taxonomy)[1] <- "label"

# Convert subset tree to dataframe and join some taxa
tree_s15_df <- tidytree::as_tibble(tree_s15)
tree_s15_df <- treeio::full_join(tree_s15_df, gtdb_taxonomy[c("label", "phylum", "class")], by = "label")

# Summarise phyla
tree_s15_sum_df <- tree_s15_df %>%
  filter(!is.na(phylum)) %>%
  dplyr::group_by(phylum) %>%
  summarise(n = n()) %>%
  dplyr::mutate(phylum_n = paste0(phylum, " (", n, ")"))
tree_s15_df <- left_join(tree_s15_df, tree_s15_sum_df, by = c("phylum"))
rm(tree_s15_sum_df) #clean

@@

# Plot circular tree
#plot_tree_s15 <- ggtree(tree_s15_md, layout = "circular", size = 0.2)
# Plot tree adding phylum
#plot_tree_s15_phyla <- plot_tree_s15 +
#  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1) +
#  geom_tiplab(aes(label = phylum, color = phylum), size = 1.75, hjust = -0.1) +
#  scale_colour_discrete(breaks = c("c", "d"), na.value = NA) +
#  scale_colour_manual(values = tree_colors, na.translate = FALSE, guide = "none") +
  #guides(color = guide_legend(title = "Phyla", ncol = 1, order = 1)) +
#  theme()
#ggsave("plot_biom_MAGs_tree_phyla.pdf", plot_tree_s15_phyla, height = 5, width = 7)

# Process QC attributes to match tree tips order
tree_tips <- tree_s15$tip.label
tree_com <- vector()
tree_con <- vector()
tree_st  <- vector()
tree_gs  <- vector()
tree_gc  <- vector()
tree_pg  <- vector()
for (i in 1:nrow(tree_s15_df)) {
  for (j in 1:nrow(mags_qc)) {
    if (tree_s15_df$label[i] == mags_qc$bin[j]) {
      tree_com <- append(tree_com, mags_qc$completeness[j])
      tree_con <- append(tree_con, mags_qc$contamination[j])
      tree_st  <- append(tree_st,  mags_qc$strain_heterogeneity[j])
      tree_gs  <- append(tree_gs,  mags_qc$genome_size[j])
      tree_gc  <- append(tree_gc,  mags_qc$gc[j])
      tree_pg  <- append(tree_pg,  mags_qc$predicted_genes[j])
    }
  }
}

# Save attributes in a dataframe
tree_qc_df <- tibble(Bin = tree_tips, Completeness = tree_com,
                     Contamination = tree_con, Strain_Het = tree_st,
                     Genome_Size = tree_gs, GC = tree_gc, Pred_Genes = tree_pg)
rm(tree_tips, tree_com, tree_con, tree_st, tree_gs, tree_gc, tree_pg, i, j) #clean memory

# Plot tree adding QC as multiple layers
plot_tree_s15_md <- ggtree(tree_s15_md, layout = "circular", size = 0.1) +
  geom_tippoint(aes(color = phylum), na.rm = TRUE, size = 1) +
  geom_tiplab(aes(label = phylum, color = phylum), size = 2, hjust = -0.2) +
  geom_treescale(x = -0.15, y = 0, offset = 0.2, fontsize = 1.4, linesize = 0.1, width = 0.1) +
  scale_colour_manual(values = tree_colors, guide = "none") +
  #add QC as layers
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Completeness, fill = Completeness),
             orientation = "y", stat = "identity", position = "auto",
             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.3, offset = 1,
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3)) +
  scale_fill_gradient(name = "Completeness (%)", low = "#1f4060", high = "#4497dd",
                      guide = guide_legend(keyheight = 0.3, keywidth = 0.6, reverse = TRUE, order = 1)) +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Contamination, fill = Contamination),
             orientation = "y", stat = "identity", position = "auto",
             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.3, offset = 0.15,
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3)) +
  scale_fill_gradient(name = "Contamination (%)", low = "#6a60a8", high = "#c4c2de",
                      guide = guide_legend(keyheight = 0.3, keywidth = 0.6, reverse = TRUE, order = 2)) +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Strain_Het, fill = Strain_Het),
             orientation = "y", stat = "identity", position = "auto",
             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.3, offset = 0.15,
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 3)) +
  scale_fill_gradient(name = "Strain Heterogeneity", low = "#c9543f", high = "#e9bab2",
                      guide = guide_legend(keyheight = 0.3, keywidth = 0.6, reverse = TRUE, order = 3)) +
  new_scale_fill() + new_scale_color() +
  geom_fruit(data = tree_qc_df, geom = geom_bar,
             mapping = aes(y = Bin, x = Pred_Genes, fill = Pred_Genes),
             orientation = "y", stat = "identity", position = "auto",
             grid.params = list(linetype = 1, size = 0.1), pwidth = 0.3, offset = 0.15,
             axis.params = list(axis = "x", text.size = 1.8, vjust = 1, nbreak = 2)) +
  scale_fill_gradient(name = "Predicted Genes", low = "#279854", high = "#a9dede",
                      guide = guide_legend(keyheight = 0.3, keywidth = 0.6, reverse = TRUE, order = 4)) +
  theme(legend.title = element_text(size = 10),
        legend.text = element_text(size = 9),
        legend.key.height = unit(4, units = "mm"),
        legend.key.width = unit(3, units = "mm"))
ggsave("article_profiling/1-taxonomy/plots/plot_biom_MAGs_tree_md.pdf", plot_tree_s15_md, height = 7, width = 9)
```

@@

```{r fig1}
# Merge plots using patchwork
fig_1a <- wrap_plots(plot_tree_s15_md, plot_mags_qc, plot_depth_sgbs_biom,
                     design = "\nAAAB\nAAAC", heights = c(1, 0.85))
#ggsave("article_profiling/fig_1a.pdf", fig_1a, height = 8, width = 15)
fig_1b <- wrap_plots(plot_depth_sgbs_both, plot_depth_spec_eco, plot_depth_kggbs_biom,
                     design = "\nAAABBCC\n")
#ggsave("article_profiling/fig_1b.pdf", fig_1b, height = 8, width = 15)

fig_1 <- wrap_plots(fig_1a, fig_1b, nrow = 2, heights = c(1, 0.85)) +
  plot_annotation(tag_levels = list(c("a", "b", "c", "d", "", "e", "f"))) &
  theme(plot.tag = element_text(face = "bold"))
ggsave("article_profiling/fig_1.pdf", fig_1, height = 16, width = 16)
#ggsave("article_profiling/fig_1.png", fig_1, height = 16, width = 16)
rm(fig_1a, fig_1b) #clean

#sup 1
fig_s1 <-  { plot_depth_top_phyla |
    (plot_depth_usgbs / plot_depth_spec + plot_layout(heights = c(0.85, 0.15))) } +
  plot_layout(ncol = 2, widths = c(0.25, 1)) +
  plot_annotation(tag_levels = list(c("a", "b", "c"))) &
  theme(plot.tag = element_text(size = 20, face = "bold"))
fig_s1[[2]][[1]] <- fig_s1[[2]][[1]] + theme(axis.title.y = element_text(vjust = 0))
fig_s1[[2]][[2]] <- fig_s1[[2]][[2]] + theme(axis.line.y.left = element_blank(),
                                             axis.ticks.y.left = element_blank())#,
                                             #axis.text.y = element_text(hjust = 2.2))
ggsave("article_profiling/fig_s1.pdf", fig_s1, height = 13, width = 22)
#ggsave("article_profiling/fig_s1.png", fig_s1, height = 13, width = 22)
```

@@

# Populate both pathways and metabolism
kegg_ko_paths <- vector()
kegg_ko_metab <- vector()
for (i in 1:nrow(kegg_paths)) {
	kegg_ko_paths <- append(kegg_ko_paths, fx_setKOpathways(kegg_paths$Definition[i]))
	kegg_ko_metab <- append(kegg_ko_metab, fx_setKOmetabtypes(kegg_paths$Definition[i]))
}
kegg_paths["Pathway"] <- kegg_ko_paths
kegg_paths["Metabolism"] <- kegg_ko_metab
rm(kegg_ko_paths, kegg_ko_metab) #clean memory

@@

### Bubble and boxplots
